\documentclass{article}
\input{preamble.tex}
\pagestyle{fancy}
\lhead{Algebraic Topology Notes}
\rhead{Thomas Boyko}
\chead{}

\begin{document}
\setcounter{section}{-1}

\section{Homotopy and CW Complexes}

    \begin{defn}[Homotopy of Paths]
    $ f,g:X\to Y $ are homotopic if there exists some $ F:X\times I\to Y $, so that $ f(x)=F(x,0) $ and $ g(x)=F(x,1)  $. This induces an equivalence relation on the set of maps $ X\to Y $.
    \begin{proof} 
        Consider the following maps:
        \begin{enumerate}
            \item $ F:X\times I\to Y $, defined by 
                \[
                F(x,t)=f(x)
                .\] 
            \item Suppose $ f,g $ are homotopic by some $ F $, define $ \bar{F}:X\times I\to Y $ by 
                \[
                \bar{F}(x,t)=F(x,1-t)
                .\] 
            \item Suppose $ f\sim ^{F_1}g\sim ^{F_2} h$.
        \end{enumerate}
    \end{proof}
\end{defn}
\begin{defn}[Homotopy Equivalence]
    $ f:X\to Y $ is a Homotopy equivalence if $ \exists g:Y\to X $:
    \begin{enumerate}
        \item $ fg $ is homotopic to $ \mathbb{1}_Y$ 
        \item $ gf $ is homotopic to $ \mathbb{1}_Y $.
    \end{enumerate}
\end{defn}
\begin{defn}
    $ \iota:X\to Y $ is an inclusion of retraction if $\exists r:Y\to X $, so that $ \iota r=id_X $
\end{defn}

\begin{defn}
    $ X $ is a Hausdorff space if $ \forall p\neq q\in X $, there exist disjoint open $ U,V $ containing $ p,q $
\end{defn}
\begin{defn}
    $ X  $ is compact, if for every open cover $ \bigcup_{\alpha\in A} G_\alpha $ of $ X, $ there exists some finite subcover $ \bigcup_{i=1} ^{n}G_{\alpha_i} $.
\end{defn}

\begin{thm}
    Recall the following:
    \begin{enumerate}
        \item If $ A\subseteq X $ is a closed subset of a compact space, then $ A  $ is compact.
        \item If $ X $ is Hausdorff and $ A\subseteq X $ is compact then $ A $ is closed
        \item The image of a compact set under a continuous map is compact.
    \end{enumerate}

\end{thm}
    \begin{defn}
        Let $ X  $ a space and $ \sim   $ and equivalence relation on $ X $. Define a topology on the set of equivalence classes $ X / \sim  $, by considering $ \pi $ the projection map. Say $ \pi^{-1}(U) $ open exactly when $ U $ is open.

        For a subspace $ A \subseteq X $ we may define $ X / A $ by using the relation $ \sim _A $, by $ a\sim b $ if $ a=b $ or both $ a,b\in A $.
    \end{defn}

    \begin{lem}
        Suppose $ X $ has a n equivalence relation $\sim$, and $ f:X\to Y  $ a continuous map.
        \begin{enumerate}
            \item The map $ f/ \sim :X / \sim \to Y $ is continuous
            \item If $ X $ is compact and Hausdorff, and $ f/ \sim $ is bijective, then $ f/ \sim $ is a homeomorphism.
        \end{enumerate}
    \end{lem}

    \newpage
    \begin{exmp}
        Take the compact Hausdorff space,space $ D^{n} $,  and identify the subspace $ S^{n-1} $. Define $ a\sim b $ if $ a=b $ or $ a,b \in S^{n-1}$.

        Then define $ f: $
    \end{exmp}

    \begin{lem}
        Let $ X $ be Hausdorff and $ A\subseteq X $ compact. Then $ X / A $ is Hausdorff.
    \end{lem}
    \begin{lem}
        Let $ X $ be Hausdorff, $f:S^{n-1}\to X$ be continuous. Define $Y=X\cup_f D^{n}=X\sqcup D^{n} / S^{n-1}$. $Y$ is then Hausdorff.
    \end{lem}
    Next we discuss a couple of methods of constructing spaces which are simple to work with.
    \begin{defn}
        A Finite CW complex (cell complex) is defined inductively:
        $X^{0}\subseteq X^{1}\subseteq \dots\subseteq X^{n}$
        \begin{enumerate}
            \item $X_0$ is a discrete set.
            \item $X^1= X^{0}\sqcup D^1 / f_\alpha : S^{0}_\alpha=\partial D^{1}_\alpha \to X^{0}$ % TODO
        \end{enumerate}
    \end{defn}
    \begin{thm}
        Finite cell complexes are compact and Hausdorff.

        Note: Many geomoetric spaces (compact manifolds) are homotopy equivalent to finite CW complexes.
    \end{thm}

    \begin{defn}
        
    Let $X,p$ be a topological space with chosen basepoint $p$. We define a wedge product
    $(X,p)\lor (Y,q)=(X\sqcup Y) / p\sim q$ 
    \end{defn}

    \begin{defn}
        The smash product is defined:
        \[
            (X,p)\land(Y,q) = (X\times Y) / X\lor Y
        .\]  
    \end{defn}
    \begin{defn}
        We define the join $X*Y=X\times Y\times I/\sim $ where $\sim $ identifies:
        \begin{enumerate}
            \item $(X,0,y)\sim (x',0,y)$
            \item $(x,1,y)\sim (x,1,y')$
        \end{enumerate}
    \end{defn}
    \begin{defn}
        If $X$ is a cell complex, a subcomplex of $X$ is a closed union of cells.
    \end{defn}
    \subsection{Projective Space}
    \begin{defn}
        Real Projective Space $\mathbb{R}P^{n}=S^{n} / a \sim -a$
    \end{defn}
\section{The Fundamental Group}

\section{Homology Groups}

\section{Homotopy Groups}


\end{document}
