\documentclass{article}
\input{preamble.tex}
\pagestyle{fancy}
\lhead{Assignment \# $3$}
\rhead{Name: Thomas Boyko; UCID: 30191728}
\chead{}

\begin{document}
\begin{enumerate} 

\item An $m \times n$ matrix is said to be a queen if the restriction of A to the orthogonal complement of its kernel is an isometry.

\begin{enumerate}[label= (\alph*)] 
\item Show that $A$ is a queen if and only if $A^* A$ is an orthogonal projection.
    \paragraph{Solution: } We know already that $(A^* A)^* =A^* A$, so me need only prove that $(A^* A)^2=A^* A$.

    Suppose that $A$ is a queen. Then for any $v\in (\ker A)^{\perp}=\text{ran }A^* $, $Av=v$. But since  $v\in \text{ran }A^* $, there exists some $w$ so that $A^* w=v$.



\item Show that $A$ is a queen if and only if $AA^*$ is an orthogonal projection.
    \paragraph{Solution: }

    Conversely, suppose that $A^* A$ is an orthogonal projection. Choose $v\in (\ker A)^{\perp}=\text{ran }A^{*},$ so that we have some $w$, where  $A^* w=v$.
    \begin{align*}
        \|Av\|^2&=\langle Av,Av \rangle\\
                &=\langle A^* Av,v \rangle\\
                &=\langle A^* A A^* w,A^* w \rangle\\
                &=\langle AA^* A A^* w, w \rangle\\
                &=\langle AA^*w, w \rangle\\
                &=\langle A^*w, A^* w \rangle\\
                &=\langle v,v\rangle\\
                &= \|v\|^2 \\
        \|Av\|&=\|v\|
    .\end{align*}
\item Show that a queen A is an isometry if and only if $\ker A = {0}$.

    \paragraph{Solution: }If $\ker A=\{0\} $, then $(\ker A)^{\perp}=V$, so the restriction of $A$ to the orthogonal complement of its kernel is $A$ restricted to its domain. Then $A$ is an isometry on any vector.

    Conversely, suppose $A$ is an isometry. Then:
    \begin{align*}
        v\in \ker A&\iff Av=0\\
                   &\iff \|Av\|=\|v\|=0\\
                   &\iff v=0
    .\end{align*}
    Therefore $\ker A=\{0\} $. 

\item Find an example of a $4 \times  2$ queen that has non-zero kernel. Be sure to prove it's a queen!
    \paragraph{Solution: }Take the matrix:
    \[
        A=\begin{bmatrix} 0&1\\0&0\\ 0&0\\0&0  \end{bmatrix} 
    .\] 
    Then clearly $A\begin{bmatrix} t\\0 \end{bmatrix} =0$ for any $t\in \mathbb{C}$, so $A$ has nonzero kernel, and all we must show is that $A$ is a queen.

    Begin by observing that since $\ker A=\text{span} \left\{ \begin{bmatrix} 1\\0 \end{bmatrix}  \right\} $
    , we can find:

    $\left( \ker A \right) ^{\perp}=\text{span}\left\{ \begin{bmatrix} 0\\1 \end{bmatrix}  \right\}  $. Then let $v=\begin{bmatrix} 0\\t \end{bmatrix} \in (\ker A)^{\perp}$. Computing both $\|v\|,\|Av\|,$ we see:
    \[
    \|Av\|=\Vert \begin{bmatrix} 0&1\\0&0\\ 0&0\\0&0  \end{bmatrix} \begin{bmatrix} 0\\t \end{bmatrix} \Vert=\|\begin{bmatrix} t\\0\\0\\0  \end{bmatrix} \|=|t|=\|\begin{bmatrix} 0\\t \end{bmatrix} \|=\|v\|
    .\] 
    So the restriction of $A$ to the orthogonal complement of its kernel is an isometry, and $A$ is a queen.
\end{enumerate}
\item \begin{enumerate}[label= (\alph*)] 
\item Given a singular value decomposition $A = W \Sigma V^* $ of a square matrix $A$, construct a polar decomposition of $A$ using $W,V,\Sigma$.

\item Using the method above, compute a polar decomposition for
\[
    A=\begin{bmatrix} 1&3\\2&6 \end{bmatrix} 
.\] 
\end{enumerate}

\item Find your favorite $4 \times  2$ matrix A of rank 2 and compute a singular value decomposition for A. All of the entries of A must be nonzero.

\item For an $m \times  n$ matrix $A$, show that the set of nonzero eigenvalues for $A^*  A$ coincide with that of $AA^* $.

\item Suppose $A = W \Sigma V^* $ is a singular value decomposition for $A$. Show that the columns of $W$ are eigenvectors for $AA^* $.

\end{enumerate}
\end{document}
