\documentclass{article}
\input{preamble.tex}
\usepackage[all]{xy}
\pagestyle{fancy}
\lhead{Week \# $3$}
\rhead{Name: Thomas Boyko; UCID: 30191728}
\chead{}

\begin{document}
\begin{enumerate} 
    \item Find three examples of adjoint functors not mentioned above. Do the same for initial and terminal objects.
        \begin{enumerate}
            \item Let $\mathbb{E}$ be a field extension of another field $\mathbb{F}$. Let $G$ be the Galois group $G=\mathrm{Gal}(\mathbb{E}:\mathbb{F})$ and take the category $\mathcal{E}$, with objects being intermediate field extensions between $\mathbb{E}$ and $\mathbb{F}$, and morphisms as inclusions. Take $\mathcal{G}$, with objects being subgroups, and morphisms $\mathrm{Hom}(H,H')=\{\leq |H\leq H'\} $.
                
                Then, for $\mathbb{k}\in \mathcal{E}$, we define the map:
                \[
                \phi:\mathcal{E}\to \mathcal{G}^{op},\quad\phi(\mathbb{k})= \mathrm{Gal}(\mathbb{E}:\mathbb{k})
                .\] 
                And for $H\in \mathcal{G}$, define:
                \[
                \phi:\mathcal{G}^{op}\to \mathcal{E},\quad\psi(H)=\{x\in \mathbb{E}:\alpha(x)=x\forall \alpha\in H\} 
                .\] 
                (Recalling that elements of $H$ are automorphisms of $\mathbb{E}$).
                A study of Galois theory tells us that these maps produce a bijective correspondence of field extensions and subgroups, we see that they are functors which are both adjoint to each other.
            \item In the category of fields of characteristic  $p$, for some fixed $p$, the field $\mathbb{F}_p$ is initial.
            \item 
        \end{enumerate}
    \item 2.1.14 (may be hard)

    \item Let $G$ be a group.
        \begin{enumerate}
            \item What interesting functors are there (in either direction) between $\mathrm{Set}$ and the category $[G, \mathrm{Set} ]$ of left $G$-sets? Which of those functors are adjoint to which?
                \paragraph{Solution: } Consider $F$, the forgetful functor from $[G,\mathrm{Set}]\to \mathrm{Set}$, which takes an action and maps it to the set it acts on.

                Then take $L:\mathrm{Set}\to [G,\mathrm{Set}]$, which takes a set $X$ in $\mathrm{Set}$ and maps it to an action $G\times X$, defined by $h(g,x)=(hg,x)$. For a functor in $\mathrm{Set}$, we map $L(f(x))=(g,f(x))$

                No take $R:\mathrm{Set}\to [G,\mathrm{Set}]$ defined by $R(X)=\mathrm{Map}(G,X)=\{f:G\to X\} $. We equip $\mathrm{Map}(G,X)$ with an action, $g\phi(h)=\phi(gh)$. For a functor $g$ in $\mathrm{Set}$, we map $R(g)=$

                Finally we claim that these are adjoint:
                \[\xymatrix{
                        [G,\mathrm{Set}]\ar[d]|F\\
                        \mathrm{Set}\ar@<2.5ex>[u]_{\dashv}^{L}\ar@<-2.5ex>[u]^{\dashv}_{R}
                }\]
        \end{enumerate}
    \item Let $f:A\to B$ and $g:B\to A$ be order-preserving operations between ordered sets. Prove TFAE:
        \begin{enumerate}
            \item For any $a\in A,b\in B$, 
                \[
                f(a)\leq b\iff a\leq f(b)
                .\] 
            \item $a\leq g(f(a))$ for any $a\in A$ and $f(g(b))\leq b$ for any $b\in B$
        \end{enumerate}
    \item Show that for any adjunction, the right adjoint is full and faithful if and only if the counit is an isomorphism
\end{enumerate}
\end{document}
